Heuristics for the Score-Constrained Strip-Packing Problem

### 2018-07-06 PAPER/main.tex
rearranged locations of figures and tables, minor changes

During MCM, if a vertex $v_i$ is not adjacent to any other unmatched vertex except its partner $p(v_i)$ via a blue edge, the preceeding vertex $v_{i-1}$ can be ``rematched'', provided that 
\begin{enumerate*}[label={(\alph*)}]
	\item $v_i$ is not the first vertex; 
	%\item $v_{i-1} \in R'$; and
	%\item $(v_{i-1}, m(v_{i-1})) \in R'$; and
	\item the previous vertex $v_{i-1}$ has been matched; and
	%\item $(v_{i-1}, p(v_i)) \in R$.
	\item $v_{i-1}$ and $p(v_i)$ are adjacent via a red edge in $R$.
\end{enumerate*}
Then, $v_i$ is matched with the vertex that is currently matched with $v_{i-1}$, and $v_{i-1}$ is rematched with $p(v_i)$. It is guaranteed that MCM will produce a maximum cardinality matching $R' \subseteq R$, and is able to do so in at most $O(n \log n)$ time (\citealp{mahadev1994}).

@@

BR operates by first ordering the edges in $R'$ such that the lower-indexed vertices of each edge are in increasing order and the higher-indexed vertices are in decreasing order, i.e. $(v_1, v_{2n+2}), (v_2, v_{2n+1}), ..., (v_{n+1}, v_{n+2})$. Any edges that cannot be placed in such an order are then removed from this list. For instance, in the example illustrated in Figure \ref{fig:matching/mps}, the edges would be sorted as follows: $(v_1, v_{14}), (v_2, v_{13}),(v_3, v_{12}),(v_4, v_{11}),(v_6, v_{10}),(v_7, v_8)$. Note that, since $v_5$ was not matched with the highest-indexed vertex possible during MCM, the edge $(v_5, v_9)$ does not adhere to the required structure of the list and is therefore omitted.
% (v5, v9) removed from list, not matched with highest-indexed vertex in MCM since it was v10 and v10 is v5's partner.

Starting from the first edge in the list, BR then searches through the list to find an edge that meets the following conditions: 
\begin{enumerate*}[label={(\alph*)}]
	\item the lower-indexed vertex of the current edge and the higher-indexed vertex of the next edge in the list are adjacent via an edge in $R$; and
	\item the current edge and the next edge are members of different components on $G'$.
\end{enumerate*}
If these conditions are met, BR adds the current edge to a set $R_i$, and continues to add all succeeding edges in the list to $R_i$ provided that, for each edge, both conditions hold and the succeeding edge is not a member of a component of $G'$ that has an edge in $R_i$. Once there are no more valid edges available to add to $R_i$, BR resumes its search through the remaining edges in the list to find a new edge that meets the conditions, and can begin a new set $R_{i+1}$. The procedure terminates once the penultimate edge in the list has been assessed.
% (a) (v_i,m(v_{i+1})) \in R
% (b) (v_i, m(v_i)) \in C_k, (v_{i+1}, m(v_{i+1})) \in C_l, k \neq l

@@

\item Otherwise, it may be that multiple sets are required to connect the components of $G'$. For two sets $R_i$ and $R_j$ to ``overlap'', each set must have one edge from the same component in $G'$, with the other edges in each set being from different components. For example, if $R_1$ has two edges, one from component $C_1$ and one from component $C_2$, and $R_2$ has three edges that are members of components $C_2$, $C_3$ and $C_4$, then $R_1$ and $R_2$ are said to overlap, since they each possess an edge $C_2$, but the other edges in the sets are from different components. Hence, a selection of sets need to be found such that each set overlaps with at least one other set, and each component has at least one edge in one of the selected sets. If such a collection of sets exists, then the components of $G$ can be merged to create an alternating Hamiltonian cycle by applying the connecting procedure above to every set in the collection.
% Refer to figures
\end{enumerate}
Using the instance illustrated in the figures as an example, the edges $(v_3, v_{12})$ and $(v_4, v_{11})$ are in the set $R_1$ formed by BR, as \begin{enumerate*}[label={(\alph*)}]
	\item $(v_3, v_{11}) \in R$, i.e. the lower-indexed vertex of the first edge is adjacent to the higher-indexed vertex of the next edge; and
	\item the edges are members of different components ($(v_3, v_{12}) \in C_1$ and $(v_4, v_{11}) \in C_2$ as seen in Figure \ref{fig:mps}).
\end{enumerate*}
Note that since $v_4$ and $v_{11}$ are adjacent, $v_4$ must also be adjacent to $v_{12}$, i.e. $(v_4, v_{12}) \in R$, since the value associated with $v_{12}$ is greater than or equal to the value associated with $v_{11}$. Then, as $|R_1| = q$, the edges $(v_3, v_{12})$ and $(v_4, v_{11})$ are removed from $R'$ and replaced by the edges $(v_3, v_{11})$ and $(v_4, v_{12})$ from $R$, as shown in Figure \ref{fig:connect/cycle}. Removing the dominating vertices and any incident edges results in an alternating Hamiltonian path, which corresponds to a feasible solution $\mathcal{T}$.

@@

\begin{figure}	
	\centering
	\begin{subfigure}[H]{0.57\textwidth}
		\includestandalone[width=\textwidth]{figures/mpsconnect}
		\caption{Edges from $R\backslash R'$ can connect vertices from different cycles.}	
		\label{fig:mpsconnect} \vspace{5mm}
	\end{subfigure}
	\begin{subfigure}[H]{0.57\textwidth}
		\includestandalone[width=\textwidth]{figures/mpscycle}
		\caption{Replacing edges connects the components and forms an alternating Hamiltonian cycle.}	
		\label{fig:mpscycle} \vspace{5mm}
	\end{subfigure}
	\begin{subfigure}[H]{0.7\textwidth}
		\includestandalone[width=\textwidth]{figures/mpspath}
		\caption{Alternating Hamiltonian Path corresponding to feasible solution $\mathcal{T}$.}	
		\label{fig:mpspath}
	\end{subfigure}
	\caption{Selected edges found using BR merge the components of $G'$ together, forming an alternating Hamiltonian cycle which corresponds to a feasible solution $\mathcal{T} = \langle(1,2), (6,3), (5,3), (4,4), (4,2), (7,1) \rangle$.}
	\label{fig:connect/cycle}
\end{figure}

\begin{comment}
\begin{algorithm}[h]
\caption{\textsc{Alternating Hamiltonian Construction} (AHC) algorithm}
\label{alg:ahca}
\begin{algorithmic}[1]
	\State Create the graph $G = (V, B \cup R)$ from instance $\mathcal{M}$
	\State Execute \textsc{Maximum Cardinality Matching}
	\If{$|R'| < n + 1$}
		\State infeasible, end
	\EndIf
	\State Produce the subgraph $G' = (V, B \cup R')$ consisting of $q$ cycles
	\If{$q = 1$}
		\State remove dominating vertices and incident edges
		\Return $G'$
	\Else
		\State execute \textsc{Bridge Recognition}
		\If{$q = 1$}
			\State remove dominating vertices and incident edges
			\Return $G'$
		\Else
			\State infeasible, end	
		\EndIf	
	\EndIf
\end{algorithmic}	
\end{algorithm}
\end{comment}

@@

In this section, we now turn our attention to the multi-strip version of the problem. As mentioned in the introduction, this problem is a generalisation of the BPP in that we also require the sum of every pair of adjacent score widths to be greater than or equal to a minimum scoring distance $\tau$. It follows that the SCSPP is at least as hard as the BPP, which is known to be NP-hard (\citealp{garey1979}), and so (under the assumption that $P \neq NP$) there is no known algorithm that is able to find an optimal solution for every instance of the SCSPP in polynomial time. Instead, heuristics can be used to find near-optimal solutions in a shorter amount of time.

@@

\begin{figure}[h!]	
	\centering
	\includestandalone[width=\textwidth]{figures/strips}
	\caption{Example instance of the sub-problem, with $W = 20$ and $\tau = 70$. In MFFD the constraint is not fulfilled in either orientation, however MFFD$^+$ is able find a feasible arrangement.}	
	\label{fig:comparestrips}
\end{figure}

>>>>

During MCM, if a vertex $v_i$ is not adjacent to any other unmatched vertex except its partner $p(v_i)$ via a blue edge, the preceeding vertex $v_{i-1}$ can be ``rematched'', provided that 
\begin{enumerate*}[label={(\alph*)}]
	\item $v_i$ is not the first vertex; 
	\item the previous vertex $v_{i-1}$ has been matched; and
	\item $v_{i-1}$ and $p(v_i)$ are adjacent via a red edge in $R$.
\end{enumerate*}
Then, $v_i$ is matched with the vertex that is currently matched with $v_{i-1}$, and $v_{i-1}$ is rematched with $p(v_i)$. It is guaranteed that MCM will produce a maximum cardinality matching $R' \subseteq R$, and is able to do so in at most $O(n \log n)$ time (\citealp{mahadev1994}).

@@

BR operates by first ordering the edges in $R'$ such that the lower-indexed vertices of each edge are in increasing order and the higher-indexed vertices are in decreasing order, i.e. $(v_1, v_{2n+2}), (v_2, v_{2n+1}), ..., (v_{n+1}, v_{n+2})$. Any edges that cannot be placed in such an order are then removed from this list. For instance, in the example illustrated in Figure \ref{fig:matching/mps}, the edges would be sorted as follows: $(v_1, v_{14}), (v_2, v_{13}),(v_3, v_{12}),(v_4, v_{11}),(v_6, v_{10}),(v_7, v_8)$. Note that, since $v_5$ was not matched with the highest-indexed vertex possible during MCM, the edge $(v_5, v_9)$ does not adhere to the required structure of the list and is therefore omitted.

Starting from the first edge in the list, BR then searches through the list to find an edge that meets the following conditions: 
\begin{enumerate*}[label={(\alph*)}]
	\item the lower-indexed vertex of the current edge and the higher-indexed vertex of the next edge in the list are adjacent via an edge in $R$; and
	\item the current edge and the next edge are members of different components on $G'$.
\end{enumerate*}
If these conditions are met, BR adds the current edge to a set $R_i$, and continues to add all succeeding edges in the list to $R_i$ provided that, for each edge, both conditions hold and the succeeding edge is not a member of a component of $G'$ that has an edge in $R_i$. Once there are no more valid edges available to add to $R_i$, BR resumes its search through the remaining edges in the list to find a new edge that meets the conditions, and can begin a new set $R_{i+1}$. The procedure terminates once the penultimate edge in the list has been assessed.

@@

\item Otherwise, it may be that multiple sets are required to connect the components of $G'$. For two sets $R_i$ and $R_j$ to ``overlap'', each set must have one edge from the same component in $G'$, with the other edges in each set being from different components. For example, if $R_1$ has two edges, one from component $C_1$ and one from component $C_2$, and $R_2$ has three edges that are members of components $C_2$, $C_3$ and $C_4$, then $R_1$ and $R_2$ are said to overlap, since they each possess an edge $C_2$, but the other edges in the sets are from different components. Hence, a selection of sets need to be found such that each set overlaps with at least one other set, and each component has at least one edge in one of the selected sets. If such a collection of sets exists, then the components of $G$ can be merged to create an alternating Hamiltonian cycle by applying the connecting procedure above to every set in the collection.
\end{enumerate}

Using the instance illustrated in the figures as an example, the edges $(v_3, v_{12})$ and $(v_4, v_{11})$ are in the set $R_1$ formed by BR, as \begin{enumerate*}[label={(\alph*)}]
	\item $(v_3, v_{11}) \in R$, i.e. the lower-indexed vertex of the first edge is adjacent to the higher-indexed vertex of the next edge; and
	\item the edges are members of different components ($(v_3, v_{12}) \in C_1$ and $(v_4, v_{11}) \in C_2$ as seen in Figure \ref{fig:mps}).
\end{enumerate*}
Note that since $v_4$ and $v_{11}$ are adjacent, $v_4$ must also be adjacent to $v_{12}$, i.e. $(v_4, v_{12}) \in R$, since the value associated with $v_{12}$ is greater than or equal to the value associated with $v_{11}$. Then, as $|R_1| = q$, the edges $(v_3, v_{12})$ and $(v_4, v_{11})$ are removed from $R'$ and replaced by the edges $(v_3, v_{11})$ and $(v_4, v_{12})$ from $R$, as shown in Figure \ref{fig:connect/cycle}. Removing the dominating vertices and any incident edges results in an alternating Hamiltonian path, which corresponds to a feasible solution $\mathcal{T}$.

\begin{figure}	
	\centering
	\begin{subfigure}[H]{0.57\textwidth}
		\includestandalone[width=\textwidth]{figures/mpsconnect}
		%\caption{Edges from $R\backslash R'$ can connect vertices from different cycles.}	
		\label{fig:mpsconnect}
	\end{subfigure}
	\begin{subfigure}[H]{0.57\textwidth}
		\includestandalone[width=\textwidth]{figures/mpscycle}
		%\caption{Replacing edges connects the components and forms an alternating Hamiltonian cycle.}	
		\label{fig:mpscycle}
	\end{subfigure}
	\begin{subfigure}[H]{0.6\textwidth}
		\includestandalone[width=\textwidth]{figures/mpspath}
		%\caption{Alternating Hamiltonian Path corresponding to feasible solution $\mathcal{T}$.}	
		\label{fig:mpspath}
	\end{subfigure}
	\caption{Selected edges found using BR merge the components of $G'$ together, forming an alternating Hamiltonian cycle which corresponds to a feasible solution $\mathcal{T} = \langle(1,2), (6,3), (5,3), (4,4), (4,2), (7,1) \rangle$.}
	\label{fig:connect/cycle}
\end{figure}

@@

@@

In this section, we now turn our attention to the multi-strip version of the problem. As mentioned in the introduction, the SCSPP is a generalisation of the BPP in that we also require the sum of every pair of adjacent score widths to be greater than or equal to a minimum scoring distance $\tau$. It follows that the SCSPP is at least as hard as the BPP, which is known to be NP-hard (\citealp{garey1979}), and so (under the assumption that $P \neq NP$) there is no known algorithm that is able to find an optimal solution for every instance of the SCSPP in polynomial time. Instead, heuristics can be used to find near-optimal solutions in a shorter amount of time.

@@

\begin{figure}[H]	
	\centering
	\includestandalone[width=\textwidth]{figures/strips}
	\caption{Example instance of the sub-problem, with $W = 20$ and $\tau = 70$. In MFFD the constraint is not fulfilled in either orientation, however MFFD$^+$ is able find a feasible arrangement.}	
	\label{fig:comparestrips}
\end{figure}
